---
title: "R Notebook"
output: html_notebook
---

Loading Libraries

```{r}
library(tidyverse)
library(skimr)

```


Loading Data

```{r}
covid <- read_csv("C:/Users/RUMIL/Desktop/APU/STAT 597 - Soeun Kim (Statistical Consulting)/COVID19 597.csv", locale = locale(encoding = "Latin1"))

covid <- read_csv("C:/Users/RumilLegaspi/Desktop/COVID19 597.csv ")
```



Taking a glimpse of the data
```{r}
glimpse(covid)
```


```{r}
covid %>% select(
current_status,
sex,
age_group,
`Race and ethnicity (combined)`,
hosp_yn,
icu_yn,
death_yn,
medcond_yn
) %>% as_factor()

```

Histograms to explore

```{r}
#Deaths by Age group
covid %>% filter(age_group != "Unknown") %>%
ggplot() +
aes(x = age_group, fill = death_yn) +
geom_bar(position = "dodge")+
scale_fill_hue(direction = 1) +
theme_minimal() +
theme(legend.position = "top") + coord_flip()


```
Notice the amount of “Missing” there are when filling out these forms. I’d be curious where most of these
“Missing” observations are coming from



```{r}
#Deaths by Age group
covid %>% filter(sex != "Unknown", sex != "Missing", current_status == "Laboratory-confirmed case") %>%
ggplot() +
aes(x = sex, fill = current_status) +
geom_bar(position = "dodge")+
scale_fill_hue(direction = 1) +
theme_minimal() +
theme(legend.position = "top") + coord_flip()

```
The amount of missing cases is curious.





Current status by race/ethnicity

```{r}

d <- covid %>%
ggplot() +
aes(x = `Race and ethnicity (combined)`, fill = current_status) +
geom_bar(position = "dodge", stat = "identity")+

  scale_fill_hue(direction = 1) +
theme_minimal() +
geom_text(aes(label=Number, position=position_dodge(width=0.9), vjust=-0.25) +
theme(legend.position = "top") + coord_flip()

d


# in the most simplest form
ggplot(data=dat, aes(x=Types, y=Number, fill=sample)) +
geom_bar(position = 'dodge', stat='identity') +
geom_text(aes(label=Number), position=position_dodge(width=0.9), vjust=-0.25)


covid %>%
  ggplot( aes(x = factor(site),
                          y = prop.table(stat(count)),
                          fill = factor(inj_location),
                          label = scales::percent(prop.table(stat(count))))) +
    geom_bar(position = "dodge") +
    geom_text(stat = 'count',
              position = position_dodge2(.05),
              vjust = -0.9,
              size = 3)
```




Death status by race/ethnicity
```{r}
covid %>%
ggplot() +
aes(x = `Race and ethnicity (combined)`, fill = death_yn) +
geom_bar(position = "dodge")+
scale_fill_hue(direction = 1) +
theme_minimal() +
theme(legend.position = "top") + coord_flip()
```







# Research question: Which demographic has been most affected by Covid?

Based on our previous plots of confirmed cases (for race.ethnicity) we know the top 3 are, White, Non-Hispanic, Hispanic/Latino, and Black, Non-Hispanic.
```{r}
#Creating dataset for covid positive only patients
race.ethn_pos_data <- covid %>% select(
  `Race and ethnicity (combined)`, current_status)  %>%
    filter(current_status == "Laboratory-confirmed case")

skim(race.ethn_pos_data)
glimpse(race.ethn_pos_data)
str(race.ethn_pos_data)
```

So if two categorical variables are associated we can consturct
Lets create a contingency table for our data containing only confirmed cases of COVID
```{r}



race.ethn_pos_boxplot <- covid %>% select(
  `Race and ethnicity (combined)`, current_status)  %>%
    filter(current_status == "Laboratory-confirmed case") %>%
  ggplot(aes(x = current_status, y = `Race and ethnicity (combined)`, fill = current_status)) +
  geom_boxplot() +
  geom_jitter(width = 0.10, h = 0.5, alpha = 0.5) +
  labs(x = "Failed [0] : Passed [1]", y = "Change In Overall Pain Level") +
  theme(legend.position = "top") +
  coord_flip()  +
    labs(title="Boxplot of Total Percent Change in Pain Post Saline",     subtitle="",
         caption="Pain levels relative to saline at needle insertion",
         x="Failed [0] : Passed [1]",
         y=" % Change in Pain")

---
title: "Case Study 2"
author: "chris"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = "")
library(tidyverse)
```


```{r}
# load raw data
covid_raw = read_csv("COVID19 597.csv")
```

```{r, include= FALSE}
# take quick look into raw data
head(covid_raw, 10)
```

```{r}
# create new dataset
covid <-  covid_raw %>% mutate(cdc_report_dt = as.Date(cdc_report_dt, "%m/%d/%y"),
                   pos_spec_dt = as.Date(pos_spec_dt, "%m/%d/%y"),
                   current_status = factor(current_status),
                   sex = factor(sex),
                   age_group = factor(age_group),
                   `Race and ethnicity (combined)` = factor(`Race and ethnicity (combined)`),
                   hosp_yn = factor(hosp_yn),
                   icu_yn = factor(icu_yn),
                   death_yn = factor(death_yn),
                   medcond_yn = factor(medcond_yn))
```


```{r}
# view summary of each variable
covid %>%
  select(current_status,
                 sex,
                 age_group,
                 `Race and ethnicity (combined)`,
                 hosp_yn,
                 icu_yn,
                 death_yn,
                 medcond_yn) %>%
  summary(maxsum=10)
```

```{r, include = FALSE}
# another function for quick view
glimpse
```

```{r}
# create dataset with only entries with confirmed cases
covid_confirmed <-  covid %>%
                        filter(current_status == "Laboratory-confirmed case") %>%
                        select(sex,
                               age_group,
                               `Race and ethnicity (combined)`,
                               hosp_yn,
                               icu_yn,
                               death_yn,
                               medcond_yn)


```



```{r}
# bar graph of confirmed cases by gender
covid_confirmed %>%
  select(sex) %>%
  ggplot(aes(sex)) +
  geom_bar() +
  theme(axis.title=element_blank()) +
  labs(title = "Cases by Gender")
```

```{r}
# bar graph of confirmed cases by age group
covid_confirmed %>%
  select(age_group) %>%
  ggplot(aes(y=age_group)) +
  geom_bar() +
  theme(axis.title=element_blank()) +
  labs(title = "Cases by Age Groups")
```

```{r}
# bar graph of confirmed cases by race and ethnicity
covid_confirmed %>%
  select(`Race and ethnicity (combined)`) %>%
  ggplot(aes(y=`Race and ethnicity (combined)`)) +
  geom_bar() +
  theme(axis.title=element_blank()) +
  labs(title = "Cases by Race and Ethnicity")
```

```{r}
covid_confirmed %>%
  filter(sex == "Female" | sex == "Male",
         death_yn == "Yes" | death_yn == "No") %>%
  ggplot(aes(death_yn)) +
  geom_bar() +
  theme(axis.title=element_blank()) +
  labs(title = "Fatal cases")
```






```{r}
covid_confirmed %>%
  filter(medcond_yn == "Yes" | medcond_yn == "No",
         sex == "Female" | sex == "Male",
         death_yn == "Yes" | death_yn == "No") %>%
  ggplot(aes(y=medcond_yn, fill = death_yn)) +
  geom_bar() +
  theme(axis.title=element_blank()) +
  labs(title = "Fatal cases with a medical condition and without medical condition")

  #summary(maxsum = 10)
 
```




---------------------------------------------------------


race.ethn_pos_boxplot




race.ethn_pos_boxplot <- ggplot(data = covid, aes(reorder(site, sal_change_in_pain, na.rm = TRUE), sal_change_in_pain))
s_19 + geom_boxplot(varwidth=T, fill="plum") +
    labs(title="Box plot",
         subtitle="Change in Pain During Saline Injection Grouped by Site 2019 \nAscending in order of median",
         caption="Source: Initial Pain - 20 min Saline Pain",
         x="Site",
         y="Saline Change In Pain")

```

```{r}

````




```
Know where 











